---
title: "Missing Person Project"
output: html_notebook
---

```{r}
library(tidyr)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggExtra)
library(lubridate)
library(tm)
library(tidytext)
library(tidyverse)

setwd("D:/rprojects/fall2020/fall2020")

adultfemale <- read.csv("D:\\rprojects\\fall2020\\fall2020\\adultfemale.csv")
adultmale <- read.csv("D:\\rprojects\\fall2020\\fall2020\\adultmale.csv")
minorsall <- read.csv("D:\\rprojects\\fall2020\\fall2020\\minorsall.csv")


```

Week One:

1)Upload missing person data set to Rstudio as data frame and Github

2) Answer Question
	A) how many missing cases are there by gender by state
		how many gender minor cases by state

	B) Are there any missing data (observations)

	C) Are there any duplicate cases

3) Read through chapter 4 and 5 from R for Data Science

(Optional) Data science pipe line: Data wrangling section

```{r}
#clean up collum names

adultfemale <- rename(adultfemale, fName = "First.Name", lName = "Last.Name", ethnicity = "Race...Ethnicity", mDate = "Date.Modified", case = "ï..Case.Number", mAge = "Missing.Age")

adultmale <- rename(adultmale, fName = "First.Name", lName = "Last.Name", ethnicity = "Race...Ethnicity", mDate = "Date.Modified", case = "ï..Case.Number", mAge = "Missing.Age")

minorsall <- rename(minorsall, fName = "First.Name", lName = "Last.Name", ethnicity = "Race...Ethnicity", mDate = "Date.Modified", case = "ï..Case.Number", mAge = "Missing.Age")


#A) how many missing person cases are there by gender by state
#   how many gender minor cases by state 

af.cntbyst <- adultfemale %>% count(State)
am.cntbyst <- adultmale %>% count(State)
ma.cntbyst <- minorsall %>% count(State, Sex)

ma.cntbyst
am.cntbyst
af.cntbyst

#B) Are there any missing data (observations)

sum(!is.na(adultfemale$case))
sum(is.na(adultmale))
sum(is.na(minorsall))

#C) Are there any duplicate cases

anyDuplicated(adultfemale$case, incomparables = FALSE)
anyDuplicated(adultmale, incomparables = FALSE)
anyDuplicated(minorsall, incomparables = FALSE)

```
Week two

  1.	Convert the date column to a date variable 

  2.	Remove years from age column

  3.	Figure out research questions and focus 

Optional: 

  1.	Clean ethnicity table 

  2.	Join tables
```{r}
#Remove years from age column

#stopwords = c(" Years", " Year", "> ", "< ")     
#x  = minorsall$Missing.Age 
#y  = adultfemale$Missing.Age
#z  = adultmale$Missing.Age

#x  =  removeWords(x,stopwords)
#y  =  removeWords(y,stopwords)
#z  =  removeWords(z,stopwords)

#minorsall$Age <- x 
#adultfemale$Age <- y
#adultmale$Age <- z

#minorsall$mAge <- as.integer()

#Convert the date column to a date variable

#minors

minorsall$DLC <- as.Date(minorsall$DLC , format = "%m/%d/%y")

minorsall <- minorsall[order(minorsall$DLC ),]

#adultfemale

adultfemale$DLC <- as.Date(adultfemale$DLC , format = "%m/%d/%y")

adultfemale <- adultfemale[order(adultfemale$DLC ),]

#adultmale

adultmale$DLC <- as.Date(adultmale$DLC , format = "%m/%d/%y")

adultmale <- adultmale[order(adultmale$DLC ),]

```

Week 3

A. Join the data frames together

B. Clean the age column. figure out why the < > symbols are having issues. 

C. Clean the ethnicity variable 

D. Read Chapters 12 & 13

optional:

Explore the R cheat Sheet
```{r}
#Join the data frames together

mpAll <- full_join(adultfemale, adultmale, minorsall, by = c("case", "DLC", "lName", "fName", "mAge", "City", "County", "State", "Sex", "ethnicity", "mDate")) 

mpAll$DLC <- as.Date(mpAll$DLC , format = "%m/%d/%y")

#Clean the age column. figure out why the < > symbols are having issues.

#fix for age column used gsub
mpAll$Age <- gsub('(\\d\\d?)(.*)','\\1',mpAll$mAge)

mpAll$Age <- as.numeric(mpAll$Age)

mpAll[is.na(mpAll$Age),]

```

Week 4

A. standardize the ethnicity column
B. work on age range issue

```{r}


levels(as.factor(mpAll$ethnicity))

mpAll$ethnicity2 <- ifelse(mpAll$ethnicity == "American Indian / Alaska Native" |
                          mpAll$ethnicity == "American Indian / Alaska Native, White / Caucasian",
                           "American Indian/Alaska Native, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "White / Caucasian", "White, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "White / Caucasian, Other", "White, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "White / Caucasian, Hispanic / Latino", "Hispanic/Latino",
                    ifelse(mpAll$ethnicity == "Black / African American", "Black, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "American Indian / Alaska Native", "American Indian / Alaska Native, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Asian", "Asian, Non-Hispanic", 
                    ifelse(mpAll$ethnicity == "Other", "Multiple/Other",
                    ifelse(mpAll$ethnicity == "Uncertain", "Multiple/Other",
                    ifelse(mpAll$ethnicity == "Asian, Hawaiian / Pacific Islander, Other", "Asian, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Hispanic / Latino, Uncertain", "Hispanic/Latino",
                    ifelse(mpAll$ethnicity == "White / Caucasian, American Indian / Alaska Native", "White, Multiple, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Hawaiian / Pacific Islander", "Native Hawaiian / Other Pacific Islander, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "White / Caucasian, Asian", "White, Multiple, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Black / African American, American Indian / Alaska Native", "Black, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "White / Caucasian, Uncertain", "White, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Hispanic / Latino, Black / African American", "Hispanic/Latino",
                    ifelse(mpAll$ethnicity == "Black / African American, Asian", "Black, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "American Indian / Alaska Native, White / Caucasian", "American Indian/Alaska Native, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "White / Caucasian, Other", "White, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Black / African American, White / Caucasian", "Black, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Hispanic / Latino, White / Caucasian", "Hispanic/Latino",
                    ifelse(mpAll$ethnicity == "White / Caucasian, Black / African American", "Black, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Black / African American, Uncertain", "Black, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Black / African American, Hispanic / Latino", "Black, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Other, Asian", "Asian, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Asian, Other", "Asian, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Black / African American, Hawaiian / Pacific Islander", "Black, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Asian, Hawaiian / Pacific Islander", "Asian, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Hispanic / Latino, American Indian / Alaska Native", "Hispanic/Latino",
                    ifelse(mpAll$ethnicity == "Asian, Black / African American", "Black, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Hispanic / Latino, Asian", "Hispanic/Latino",
                    ifelse(mpAll$ethnicity == "Asian, Hispanic / Latino", "Hispanic/Latino",
                    ifelse(mpAll$ethnicity == "", "Multiple/Other",
                    ifelse(mpAll$ethnicity == "Uncertain, Hispanic / Latino", "Hispanic/Latino",
                    ifelse(mpAll$ethnicity == "Hawaiian / Pacific Islander, Asian", "Native Hawaiian/Pacific Islander, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "White / Caucasian, Hawaiian / Pacific Islander", "Native Hawaiian/Pacific Islander, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Other, Asian, Hawaiian / Pacific Islander", "Multiple/Other, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Black / African American, Other", "Black, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Hispanic / Latino, Other", "Hispanic/Latino",
                    ifelse(mpAll$ethnicity == "Asian, White / Caucasian", "Asian, Non-Hispanic",
                    ifelse(mpAll$ethnicity == "Other, Hispanic / Latino", "Hispanic/Latino", NA))))))))))))))))))))))))))))))))))))))))))


```

```{r}


```

```{r}
nrow(mpAll$ethnicity2)

old = options(old) 
cbind(table(mpAll$ethnicity2)/nrow(mpAll$ethnicity2))

```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
