---
title: "visuals"
output: html_notebook
---

Visualization of the missing persons data

```{r}
#make factors 

library(forcats)

mpAll$City <- as.factor(mpAll$City)
mpAll$County <- as.factor(mpAll$County)
mpAll$Sex <- as.factor(mpAll$Sex)
mpAll$State <- as.factor(mpAll$State)

```

```{r}

ggplot(mpAll, aes(x = Sex)) +
  theme_bw() +
  geom_bar() +
  labs(y = "Missing People",
       title = "Missing People by Sex")

```

```{r}
#Bar Graph Ethnicity
ggplot(mpAll, aes(y = fct_rev(fct_infreq(ethnicity2)), fill = Sex)) +
  theme_bw() +
  geom_bar() +
  labs(y = "ethnicity",
       title = "Missing People by Ethnicity") +
  facet_grid(Sex~.)
```




```{r}
#Histigram by age and sex 

ggplot(mpAll, aes(x = Age, fill = Sex)) +
  theme_bw() +
  geom_histogram(binwidth = 5) +
  labs(y = "Missing Persons Count",
       x = "Age (binwidth = 5)",
       title = "Missing People by Age")

```


```{r}
ggplot(mpAll, aes(x = Age, fill = Sex)) +
  theme_bw() +
  facet_wrap(~ ethnicity2) +
  geom_histogram(binwidth = 5) +
  labs(y = "Missing Persons Count",
       x = "Age (binwidth = 5)",
       title = "Missing People by Age")
```

```{r}


ggplot(mpAll, aes(x = DLC, fill = Sex)) +
  theme_bw() +
  #facet_wrap(~ ethnicity2, scales = "free_y")+
  geom_density(alpha = 0.5) +
  labs(y = "Number of People",
       x = "Time Line",
       title = "Trend of People Missing Over Time by Sex")
```

```{r}
ggplot(mpAll, aes(y = ethnicity2, x = DLC)) +
  theme_bw() +
  geom_jitter(aes(color = Sex)) +
  geom_violin(alpha = .5) +
  labs(y = "",
       x = "",
       title = "", 
       legend = FALSE) +
  facet_grid(Sex~.)

```

```{r}

ggplot(femaDD, aes(y = iType)) +
  theme_bw() +
  geom_bar(aes(fill = dType), binwidth = 5) +
  labs(y = "Disaster Type",
       x = "Count",
       title = "FEMA Disaster")


```

```{r}

ggplot(femaDD, aes(x = state, y = iType)) +
  theme_bw() +
  geom_jitter(alpha = 0.5) +
  labs(y = "",
       x = "",
       title = "")

```

```{r}
ggplot() +
  geom_density(mpAll, mapping = aes(x = DLC, fill = Sex), alpha = .5) +
  geom_density(femaDD, mapping = aes(x = bDate), alpha = .5) +
  theme_bw() +
  #facet_wrap(~ ethnicity2, scales = "free_y")+
  labs(y = "Number of People",
       x = "Time Line",
       title = "Trend of People Missing Over Time by Sex")


```


```{r}

ggplot() +
  #geom_jitter(femaDD, mapping = aes(x = bDate, y = state))
  geom_histogram(mpAll, mapping = aes(x = DLC, color = "orange"), binwidth = 100) +
  facet_wrap(~ethnicity2) +
  theme_bw() +
  labs(y = "Number of People",
       x = "Time Line",
       title = "Trend of People Missing Over Time by Sex")

```

```{r}
ggplot() +
  geom_jitter(mpAll, mapping = aes(x = State, y = DLC)) +
  geom_jitter(femaDD, mapping = aes(y = bDate, x = state, color = dType2))+
  theme_dark() +
  labs(y = "",
       x = "",
       title = "")
```









Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
