Week One:

1)Upload missing person data set to Rstudio as data frame and Github

2) Answer Question
	A) how many missing cases are there by gender by state
		how many gender minor cases by state

	B) Are there any missing data (observations)

	C) Are there any duplicate cases

3) Read through chapter 4 and 5 from R for Data Science

(Optional) Data science pipe line: Data wrangling section

```{r}
#load in the libraries and files
library(tidyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(ggExtra)

adultfemale <- read.csv("D:\\rprojects\\fall2020\\fall2020\\adultfemale.csv")
adultmale <- read.csv("D:\\rprojects\\fall2020\\fall2020\\adultmale.csv")
minorsall <- read.csv("D:\\rprojects\\fall2020\\fall2020\\minorsall.csv")


```

```{r}

#clean up collum names

adultfemale <- rename(adultfemale, fName = "First.Name", lName = "Last.Name", ethnicity = "Race...Ethnicity", mDate = "Date.Modified", case = "ï..Case.Number", mAge = "Missing.Age")

adultmale <- rename(adultmale, fName = "First.Name", lName = "Last.Name", ethnicity = "Race...Ethnicity", mDate = "Date.Modified", case = "ï..Case.Number", mAge = "Missing.Age")

minorsall <- rename(minorsall, fName = "First.Name", lName = "Last.Name", ethnicity = "Race...Ethnicity", mDate = "Date.Modified", case = "ï..Case.Number", mAge = "Missing.Age")


#A) how many missing person cases are there by gender by state
#   how many gender minor cases by state 

af.cntbyst <- adultfemale %>% count(State)
am.cntbyst <- adultmale %>% count(State)
ma.cntbyst <- minorsall %>% count(State, Sex)

ma.cntbyst
am.cntbyst
af.cntbyst

```
```{r}
#B) Are there any missing data (observations)

sum(!is.na(adultfemale$case))
sum(is.na(adultmale))
sum(is.na(minorsall))



```

```{r}
#C) Are there any duplicate cases

anyDuplicated(adultfemale$case, incomparables = FALSE)
anyDuplicated(adultmale, incomparables = FALSE)
anyDuplicated(minorsall, incomparables = FALSE)

```

```{r}



```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
